---
title: "Assignment 1"
output: pdf_document
Autor: Group 3
---
Group 3

Names: Nastaran Rakhshan Navazkar
Anas Al-Sammarraie
Oyindenyefah Anastasia Osagie
Sophia Jamil Ahmed Chaudhary
Dharti Patel
Maureen Uzoewulu

Introduction

This report analyzes the Healthcare Diabetes dataset, focusing on patient health indicators such as glucose levels, BMI, and age. The objective is to clean, reshape, and analyze the data to uncover basic insights, compute summary statistics, and visualize relationships between variables.
```{r}
knitr::opts_chunk$set(
echo = TRUE,
message = FALSE,
warning = FALSE,
results = 'hold'
)
```
Import Dataset
print the structure of data
```{r}
data = read.csv("Healthcare-Diabetes.csv")
str(data)

```
Printing 15 first rows of data
```{r}
head(data, 15)
```
Check Column Names
```{r}
colnames(data)

```
User-Defined Functions (Mean of Variables)
## We define simple functions to calculate mean values for key variables.
```{r}
ave_age = function(data) mean(data$Age, na.rm = TRUE)
ave_age(data)

```
Filter Dataset
```{r}
library(dplyr)

# Patients over 25

filter_data = filter(data, Age > 25)
head(filter_data, 10)

# Patients with diabetes and age over 65

high_bmi_diabetic = data %>%
filter(Outcome == 1, Age > 65)
head(high_bmi_diabetic, 10)

```
Dependent & Independent Variables
```{r}
independent_var = data[, c("Pregnancies", "Glucose", "BloodPressure",
"SkinThickness", "Insulin", "BMI",
"DiabetesPedigreeFunction", "Age")]
dependent_var = data[, "Outcome"]

# Combine and convert to data frame

new_data = as.data.frame(cbind(independent_var, Outcome = dependent_var))
head(new_data, 10)


```
Remove Missing Values
```{r}
new_data = na.omit(new_data)

```
Identify and Remove Duplicates
```{r}
distinct_data = new_data %>% distinct()
head(distinct_data, 10)


```
Reorder Rows in Descending Order
```{r}
reordered_data = new_data %>%
arrange(across(everything(), ~desc(.)))
head(reordered_data, 10)
```
Rename Columns
```{r}
colnames(new_data)[colnames(new_data) == "DiabetesPedigreeFunction"] = "DP_Function"
colnames(new_data)[colnames(new_data) == "SkinThickness"] = "Skin_Thick"
colnames(new_data)[colnames(new_data) == "BloodPressure"] = "BP"
colnames(new_data)


```
Add New Variable
```{r}
new_data = new_data %>%
mutate(Glucose_Double = Glucose * 2)
head(new_data, 10)

```
Create a Training Set
```{r}
library(dplyr)
set.seed(123)

# Number of rows
n = nrow(new_data)

# Randomly select 70% of the data for training
train_indices = sample(1:n, size = 0.7 * n)

# Create training and testing sets
training_set = new_data[train_indices, ]
testing_set  = new_data[-train_indices, ]

# Check sizes
nrow(training_set)
nrow(testing_set)
```
Summary Statistics

```{r}
summary(new_data[, c("Age", "Glucose", "BMI", "BP")])
```
Statistical Measures
```{r}
mean_val = mean(new_data$Glucose, na.rm = TRUE)
median_val = median(new_data$Glucose, na.rm = TRUE)
mode_age = as.numeric(names(sort(table(new_data$Age), decreasing = TRUE)[1]))
range_age = max(new_data$Age, na.rm = TRUE) - min(new_data$Age, na.rm = TRUE)
mean_val; median_val; mode_age;range_age

```

Scatter Plot with ggplot2
```{r}
library(ggplot2)

ggplot(new_data, aes(x = Glucose, y = Age)) +
geom_point(size = 2, color = "violet") +
labs(
title = "Age vs Glucose",
x = "Glucose",
y = "Age"
) +
theme_bw()

```
Bar Plot
```{r}
library(ggplot2)
library(dplyr)

# Calculate average age by outcome
avg_age <- new_data %>%
  group_by(Outcome) %>%
  summarise(mean_age = mean(Age, na.rm = TRUE))

# Bar plot
ggplot(avg_age, aes(x = factor(Outcome), y = mean_age, fill = factor(Outcome))) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Age by Diabetes Outcome",
    x = "Outcome (0 = No Diabetes, 1 = Diabetes)",
    y = "Average Age"
  ) +
  theme_minimal()
```
Pearson Correlation

```{r}
cor_val = cor(new_data$Glucose, new_data$Age, use = "complete.obs", method = "pearson")

ggplot(new_data, aes(x = Glucose, y = Age)) +
geom_point(color = "steelblue") +
labs(
title = paste("Glucose vs Age (Pearson r =", round(cor_val, 2), ")"),
x = "Glucose",
y = "Age"
) +
theme_minimal()

```

```markdown
---
GitHub Repository: https://github.com/Nas26/Healthcare_Diabetes.git)

































